DESING AND SIMULATION SOFTWARE FOR RHODOTRON TYPE ELECTRON ACCELERATORS

### 2023-08-17 subsections/simulation/intermediate_versions/intermediate.tex
changes from Erkcan & Gokhan hocas

Magnets were modeled as major segments of a circle, defined with $\vecthreeBF{r}_{mag}$, $\textbf{R}$ and $|\textbf{B}|$. 
For the initial implementation, $\vecthreeBF{B}$ assumed to be uniform and has no leaks outside the magnet boundary (See \fromfig{is_inside_halfsphere} in \fromapp{intermediate_codes}).

@@

    \begin{subfigure}{\textwidth}
        \begin{minted}[linenos=true, autogobble, frame=lines, framesep=2mm, fontsize=\scriptsize]{c++}
            vector3d MagneticField::actOn(Electron& e){
                if (isInside(e.pos) == -1)
                    return vector3d(0,0,0);
                vector3d Bfield = getField(e.pos);                          // Calculate B vector
                vector3d F_m = (e.vel % Bfield)*eQMratio;                   // Calculate F/m vector
                vector3d acc = (F_m)/e.gamma();                             // Calculate a vector
                return acc;
            }
        \end{minted}
    \end{subfigure}
    \caption{\eEM  interaction logic from \fromeq{acc_of_E_and_B}}
    \label{fig:3D_e_EM_interaction_first}
\end{figure}
Where * and \% are, dot-product and cross-product respectably. (See \fromfig{vector3d_dot_cross_product} of \fromapp{intermediate_codes})

@@

Considering the $\vecthreeBF{E}$ is parallel to the beam path, potential difference $V$ in the trejectory until \vecNum{-0.753}{0}{0}m is
\begin{eqnarray}
    \Delta V^1 &=& - \int \dotProdBF{E}{ds} \\
             &=& - \int_{-0.753}^{0.753} -2.65616 \times dx \\
             &=& 2.65616 \times 1.506 \\
    \Delta V^1 &=& 4 MV \\
    \Delta E^1 &=& 4 MeV \\
    E_{exitTH}^1 &=& 5 MeV
\end{eqnarray}
In the second test on the other hand, the beam would be injected into a different static uniform electric field, \\
$\vecthreeBF{E}=(0, -5.31232, 0)$ MV/m where $-0.753<x<0.753$ and $-0.753<y<0.753$, $\vecthreeBF{E}=0$ elsewhere.
\begin{eqnarray}
    \Delta V^2   &=& - \int \dotProdBF{E}{ds} \\
                 &=& - \int_{0}^{0.753} -5.31232 \times dy \\
                 &=& 5.31232 \times 0.753 \\
    \Delta V^2   &=& 4 MV \\
    \Delta E^2   &=& 4 MeV \\
    E_{exitTH}^2 &=& 5 MeV
\end{eqnarray}
Therefore, in the both tests, the beam was expected to exit $\vecthreeBF{E}$ with $E_{exitTH} = 5$ MeV. 
To also measure the variance in simulation completion times, $T_{sim}$, set of $10$ runs were completed at the configuration for each $dt$ value.
\begin{figure}[H]
    \centering
    \begin{subfigure}{0.8\textwidth}
        \centering
        \includegraphics[width=\linewidth]{../../../figures/rhodoSim/statE_par.png}
        \caption*{Setup-1, $\textbf{V}_{\parallel}$=4MV}
    \end{subfigure}
    \begin{subfigure}{0.8\textwidth}
        \centering
        \includegraphics[width=\linewidth]{../../../figures/rhodoSim/statE_perp.png}
        \caption*{Setup-2, $\textbf{V}_{\perp}$=4MV}
    \end{subfigure}
    \caption{Render of the test setups. \\ $E_{in}=1$ MeV, $t_{end}=6$ns}
\end{figure}

\begin{figure}[H]
    \captionsetup[subfigure]{justification=centering}
    \captionsetup{justification=centering}
    \centering
    \begin{subfigure}{0.9\textwidth}
        \centering
        \includegraphics[width=\linewidth]{../../../figures/analiz/staticE_lf_rk2_dt-E.png}
    \end{subfigure}

    \begin{subfigure}{0.9\textwidth}
        \centering
        \includegraphics[width=\linewidth]{../../../figures/analiz/staticE_lf_rk2_dt-Tsim.png}
    \end{subfigure}
    \caption{Comparing Leap-frog, RK4-2 performance on \eE interaction\\ $E_{in}=1$MeV, $\textbf{V}_{\parallel}$=4MV, $t_{end}=6$ns}
    \label{fig:lf_rk2_par_stat_E_comparison}
\end{figure}

>>>>

Magnets were modeled as major segments of a circle, defined with $\vecthreeBF{r}_{mag}$, $\textbf{R}$ and $|\textbf{B}|$. 
For the initial implementation, $\vecthreeBF{B}$ assumed to be uniform and has no leaks outside the magnet boundary (\fromfig{is_inside_halfsphere} in \fromapp{intermediate_codes}).

@@

    \begin{subfigure}{\textwidth}
        \begin{minted}[linenos=true, autogobble, frame=lines, framesep=2mm, fontsize=\scriptsize]{c++}
            vector3d MagneticField::actOn(Electron& e){
                if (isInside(e.pos) == -1)
                    return vector3d(0,0,0);
                vector3d Bfield = getField(e.pos);                          // Calculate B vector
                vector3d F_m = (e.vel % Bfield)*eQMratio;                   // Calculate F/m vector
                vector3d acc = (F_m)/e.gamma();                             // Calculate a vector
                return acc;
            }
        \end{minted}
    \end{subfigure}
    \caption{\eEM  interaction logic from \fromeq{acc_of_E_and_B}}
    \label{fig:3D_e_EM_interaction_first}
\end{figure}
Where * and \% are, dot-product and cross-product respectably. (\fromfig{vector3d_dot_cross_product} of \fromapp{intermediate_codes})

@@

Considering the $\vecthreeBF{E}$ is parallel to the beam path, potential difference $V$ in the trejectory until \vecNum{-0.753}{0}{0}m is
\begin{eqnarray}
    \Delta V^1 &=& - \int \dotProdBF{E}{ds} \\
             &=& - \int_{-0.753}^{0.753} -2.65616 \times dx \\
             &=& 2.65616 \times 1.506 \\
    \Delta V^1 &=& 4 MV \\
    \Delta E^1 &=& 4 MeV \\
    E_{exitTH}^1 &=& 5 MeV
\end{eqnarray}
In the second test on the other hand, the beam would be injected into a different static uniform electric field, \\
$\vecthreeBF{E}=(0, -5.31232, 0)$ MV/m where $-0.753<x<0.753$ and $-0.753<y<0.753$, $\vecthreeBF{E}=0$ elsewhere.
\begin{eqnarray}
    \Delta V^2   &=& - \int \dotProdBF{E}{ds} \\
                 &=& - \int_{0}^{0.753} -5.31232 \times dy \\
                 &=& 5.31232 \times 0.753 \\
    \Delta V^2   &=& 4 MV \\
    \Delta E^2   &=& 4 MeV \\
    E_{exitTH}^2 &=& 5 MeV
\end{eqnarray}
Therefore, in the both tests, the beam was expected to exit $\vecthreeBF{E}$ with $E_{exitTH} = 5$ MeV. 
To also measure the variance in simulation completion times, $T_{sim}$, set of $10$ runs were completed at the configuration for each $dt$ value.
Results from these tests can be observed in \fromfig{lf_rk2_perp_stat_E_comparison} and \fromfig{lf_rk2_par_stat_E_comparison} of \fromapp{data_graph}.
\begin{figure}[H]
    \centering
    \begin{subfigure}{0.9\textwidth}
        \centering
        \includegraphics[width=\linewidth]{../../../figures/rhodoSim/statE_par.png}
        \caption*{Setup-1, $\textbf{V}_{\parallel}$=4MV}
    \end{subfigure}
    \begin{subfigure}{0.9\textwidth}
        \centering
        \includegraphics[width=\linewidth]{../../../figures/rhodoSim/statE_perp.png}
        \caption*{Setup-2, $\textbf{V}_{\perp}$=4MV}
    \end{subfigure}
    \caption{Render of the test setups. \\ $E_{in}=1$ MeV, $t_{end}=6$ns}
\end{figure}

### 2023-08-17 subsections/simulation/poc/poc.tex
changes from Erkcan & Gokhan hocas

%%%%%%
% Start of poc.tex
\section{Proof of Concept}
Once the need for a custom built tool for designing and simulating a Rhodotron type accelerator was apparent, a simple 1D proof of concept was implemented.

This proof of concept worked with interacting an electron with the EM field provided, at discrete time steps $dt$. The core sequence of this software is provided below.
Where $\vecthreeBF{r}$ is the position, $\vecthreeBF{v}$ is the velocity of the $e^-$;
\begin{itemize}
    \item Get electron energy $E_{in}$
    \item Get the RF field definitions
    \item Get the magnet design parameters if there is any
    \item Do following until the simulation time has ellapsed
    \begin{enumerate}
        \item Calculate $\vecthreeBF{E}_{||}$($\vecthreeBF{r}$)
        \item Calculate $\vecthreeBF{a}_{E ||}$ using \fromeq{para_and_perp_acc_of_lorentz_force}
        \item Calculate $\vecthreeBF{r}$($t+dt$) using \fromeq{leapfrog_sync_x}
        \item Calculate $\vecthreeBF{v}$($t+dt$) using \fromeq{leapfrog_sync_v}
        \item Calculate new $E_t$
    \end{enumerate}
\end{itemize}
The implemenation of this logic can be observed in \fromfig{POC_core_logic}. 
\begin{figure}[H]
    \begin{minted}[linenos=true, autogobble, frame=lines, framesep=2mm, fontsize=\footnotesize]{c++}
        for (double t=0; t<SimuTime; t+=dT){
            double RelBeta  = v/c;
            double RelGamma = 1.0 / sqrt(1.0-RelBeta*RelBeta);

>>>>

%%%%%%
% Start of poc.tex
\section{Proof of Concept}
The POC worked with interacting an electron with the EM field provided, at discrete time steps $dt$. The core sequence of this software is provided below.
Where $\vecthreeBF{r}$ is the position, $\vecthreeBF{v}$ is the velocity of the $e^-$;
\begin{itemize}
    \item Get electron energy $E_{in}$
    \item Get the RF field definitions
    \item Get the magnet design parameters if there is any
    \item Do following until the simulation time has ellapsed
    \begin{enumerate}
        \item Calculate $\vecthreeBF{E}_{||}$($\vecthreeBF{r}$)
        \item Calculate $\vecthreeBF{a}_{E ||}$ using \fromeq{para_and_perp_acc_of_lorentz_force}
        \item Calculate $\vecthreeBF{r}$($t+dt$) using \fromeq{leapfrog_sync_x}
        \item Calculate $\vecthreeBF{v}$($t+dt$) using \fromeq{leapfrog_sync_v}
        \item Calculate new $E_t$
    \end{enumerate}
\end{itemize}
The implemenation of this logic can be observed in \fromfig{POC_core_logic}. 
\begin{figure}[H]
    \begin{minted}[linenos=true, autogobble, frame=lines, framesep=2mm, fontsize=\footnotesize]{c++}
        for (double t=0; t<SimuTime; t+=dT){
            double RelBeta  = v/c;
            double RelGamma = 1.0 / sqrt(1.0-RelBeta*RelBeta);

### 2023-08-17 subsections/theory/common/common_start_theory.tex
changes from Erkcan & Gokhan hocas

\pagenumbering{arabic}

%#####
% START OF THEORY

\newcommand{\fromeq}[1]{\textit{equation \ref{eq:#1}}}
\newcommand{\fromeqs}[2]{\textit{equations \ref{eq:#1} and \ref{eq:#2}}}

@@

\chapter{THEORY}

>>>>

\pagenumbering{arabic}
\newcommand{\fromeq}[1]{\textit{equation \ref{eq:#1}}}
\newcommand{\fromeqs}[2]{\textit{equations \ref{eq:#1} and \ref{eq:#2}}}

@@

%#####
% START OF INTRODUCTION AND THEORY

\chapter{INTRODUCTION AND THEORY}

Particle accelerators are essential tools for scientific research. They enable scientists to probe the fundamental constituents of matter, study particle interactions, and explore the laws of physics. 
Accelerators have been instrumental in discovering and characterizing fundamental particles, such as quarks, leptons, and more recently the Higgs boson. 
They also facilitate the production of high-intensity particle beams for applications in industrial processes. A conservative estimate puts the number of industrial high-current electron
accelerators to be in the thousands, worldwide.

MeV-level electron beams, both as is and as X-Ray sources, are especially useful for industrial and medical applications. 
For example, high-energy beams (5–10 MeV) are used in medical accelerators and provide services such as medical device sterilization and food irradiation;
with medium energy devices (400 keV–5 MeV) wires, cables and tires are improved, factory flue gases or wastewater can be treated.
Although most of the industrial electron accelerators operating at MeV-level are linear accelerators at the moment, due to their compact nature, circular accelerators have been gaining prominence.

Rose-pattle accelerator is a type of circular accelerator suitable for producing electron beams at MeV-level energy. 
It that was developed in 1989 and was named \textit{Rhodotron} \cite{rhodo_pottier}. This name was trademarked by \textit{IBA Group} in the following years.
